Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2014-01-23 Paper/gammaRays.tex
Drafted filtering with gtselect and gtmktime

\usepackage{commath}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{subcaption}
\usepackage{units}
\usepackage{hyperref}
\usepackage{siunitx}

@@

\subsection{Burst Data Queries}

@@

Finally, to download the data about observed photons, we fill the form at the LAT data server front page with the following values:
\begin{itemize}
	\item{{\bf Object name or coordinates}. Here the coordinates are filled from the table \ref{tab:bursts}.}
	\item{{\bf Coordinate system}. We use J2000 here.}
	\item{{\bf Search radius (degrees)}. The value of $60$ is used here. Events will be filtered precisely later on by using the point spread functions (PSFs) of the LAT.}
	\item{
		{\bf Observation dates}. We extend the duration of the burst by $15\%$ relative to the table \ref{tab:bursts} values to both past and future to have a decent safety margin. So the time ranges we fill in here are:
		\begin{align*}
			\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
			\texttt{time} + \texttt{endOffset} &+ 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right)
		\end{align*}
	}
	\item{{\bf Time system}. MET is used.}
	\item{{\bf Energy range (MeV)}. The whole energy range is required for our study. So we fill in $30,\,300000$.}
	\item{{\bf LAT data type}. Extended, because we need transient photons for better statistics.}
	\item{{\bf Spacecraft data}. Checked. We need it for both events filtering, and calculation of the PSFs and exposure maps.}
\end{itemize}

The data we downloaded contains a lot of background and radiation from other sources. We need to filter it before using it.

\subsection{Data Filtering}
\subsection{Point Spread Functions}
\subsection{Exposure Maps}
\subsection{Linear Component (background)}
\subsection{KS-test}
\subsection{Results}

>>>>

\usepackage{commath}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{subcaption}
\usepackage{units}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{listings}

\lstset{columns=fullflexible,basicstyle=\ttfamily}

@@

\subsection{Burst Data Download}

@@

Finally, to download the data about observed photons, we fill the form at the LAT data server front page with the following values:
\begin{itemize}
	\item{{\bf Object name or coordinates}. Here the coordinates are filled from the table \ref{tab:bursts}.}
	\item{{\bf Coordinate system}. We use J2000 here.}
	\item{{\bf Search radius (degrees)}. The value of $60$ is used here. Events will be filtered precisely later on by using the point spread functions (PSFs) of the LAT.}
	\item{
		{\bf Observation dates}. We extend the duration of the burst by $15\%$ relative to the table \ref{tab:bursts} values to both past and future to have a decent safety margin. So the time range we fill in here is:
		\begin{align*}
			\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
			\texttt{time} + \texttt{endOffset} &+ 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right)
		\end{align*}
	}
	\item{{\bf Time system}. MET is used.}
	\item{{\bf Energy range (MeV)}. The whole energy range is required for our study. So we fill in $30,\,300000$.}
	\item{{\bf LAT data type}. Extended, because we need transient photons for better statistics.}
	\item{{\bf Spacecraft data}. Checked. We need it for both events filtering, and calculation of the PSFs and exposure maps.}
\end{itemize}

For background estimation purposes, we also download observational data during a day before the burst, specifically in the time range:
\begin{align*}
	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right) - \unit[86400]{sec},\\
	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right)
\end{align*}
All other values filled in the form are the same.

Before reading the data, we need to apply the basic filtering as described in Fermi guidelines. Two tools from the Fermi Science Tools package are used for that task (the same filtering is applied to both burst and background data).

The first one is \texttt{gtselect}. It filters the Earth limb emission. We run the tool with the following parameters\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/explore_latdata_burst.html}}:
\begin{lstlisting}
gtselect infile=[eventFile] outfile=filtered.fits
	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
	zmax=100 evclass=0 convtype=-1 evtable=EVENTS
\end{lstlisting}
Here \texttt{[eventFile]} is the file downloaded in the previous step.

The second one is \texttt{gtmktime}. It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational. Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on. Parameters are the following\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}}:
\begin{lstlisting}
gtmktime scfile=[spacecraft] evfile=filtered.fits outfile=timed.fits
	filter="DATA_QUAL>0 && LAT_CONFIG==1" roicut=yes
\end{lstlisting}
where \texttt{[spacecraft]} is the FITS containing the spacecraft data.

Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky. Before using the data obtained from it, we should perform a more elaborate filtering by location using Fermi LAT's point spread function, which will be described in the following section.

\subsection{Point Spread Functions}
\subsection{Exposure Maps}
\subsection{KS-test}
\subsection{Results}
